---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# chicri

<!-- badges: start -->
[![R-CMD-check](https://github.com/codiewood/chicago_crime/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/codiewood/chicago_crime/actions/workflows/check-standard.yaml)
<!-- badges: end -->

The goal of chicri is to provide functions for the analysis of the [Chicago crime data set](https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-Present/ijzp-q8t2). Namely, the functions are part of the SC1/SM1 group project for the Compass PhD, and aid in performing regressions and classifications.

## Installation

You can install the development version of chicri from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("codiewood/chicago_crime/chicri")
```

## Example

This is a basic example which shows you how to load the Chicago crime data from the year 2019:

```{r example}
library(chicri)
data <- load_crimes_API(year="2019")
```

We can then also process the data using the data processing functionality, and investigate the outcome.

```{r cars}
summary(data)
data <- process_data(data)
summary(data)
```

